MCMC Redistricting

### 2020-12-14 paper.Rmd
cleanup

  Fair elections are an integral part of democracy for, without it, the will of the people might not be respected and if that happens we no longer have a democracy. However, despite the huge importance of the elections in the US, there are still ways that a party can negatively impact the election by gaining an unfair advantage. One of the ways that this can happen, is gerrymandering were to gain a political advantage in a state by manipulating district boundaries. This issue has come before the supreme court, however, it was ruled that it was up to Congress and the states to deal with it.

  Due to the scale of this issue, the paper by Fifield et al, proposes the use of a Markov Chain Monte Carlo method to automate the redistricting process, despite having a few Monte Carlo algorithms existing none of them have a strong theoretical background. Fifield et al algorithm proposes to solve the automation redistricting by formulating it as a Graph Cut problem, where the MCMC samples the districts plans from arbitrary distributions over the set of n contiguous districts, they also implemented equal population constraint, which is a main requirement in the real-world problem.

## What is Redistricting?
 "In the United States is the process of drawing electoral district boundaries. A congressional act passed in 1967 requires that representatives be elected from single-member districts, except when a state has a single representative, in which case one state-wide at-large election be held." (Wikipedia). Redistricting also has to follow 6 different criteria, which are as follows.

Redistricting criteria:
    - compactness
    - contiguity
    - equal population
    - preservation of existing political communities
    - partisan fairness
    - racial fairness

@@

## Metropolis-Hastings Iteration Example
To simulate redistricting, “perturbation” looks like swapping precincts along existing district borders. \ref{fig:MCMC example} illustrates the perturbation process for an example sample in a Metropolis-Hastings chain. Perturbation works by identifying nodes that exist on congressional district borders and proposing swaps along the borders. This produces a perturbation of the previous sample which is accepted with acceptance ratio

@@

![MCMC example](figures/mcmc_example.png)

@@

>>>>

  Fair elections are an integral part of democracy for, without it, the will of the people might not be respected and if that happens we no longer have a democracy. However, despite the huge importance of the elections in the US, there are still ways that a party can negatively impact the election by gaining an unfair advantage. One of the ways that this can happen, is gerrymandering. This is the practice for unfairly obtaining a political advantage in a state by manipulating district boundaries. This issue has come before the supreme court, however, it was ruled that it was up to Congress and the states to deal with it.

  Due to the scale of this issue, the paper by Fifield et al, proposes the use of a Markov Chain Monte Carlo method to automate the redistricting process. Monte Carlo algorithms for redistricting have existed previously, but none of them have a strong theoretical background. Fifield et al algorithm proposes to solve the automation redistricting by formulating it as a Graph Cut problem, which allows the use of the Metropolis-Hastings algorthim to generate MCMC samples of districting plans from arbitrary distributions over the set of $n$ contiguous districts. They also implemented equal population constraint, which is a main requirement in the real-world problem.

## What is Redistricting?
 "In the United States is the process of drawing electoral district boundaries. A congressional act passed in 1967 requires that representatives be elected from single-member districts, except when a state has a single representative, in which case one state-wide at-large election be held." (Wikipedia). Redistricting also has to follow 6 different criteria.

    - Compactness
    - Contiguity
    - Equal population
    - Preservation of existing political communities
    - Partisan fairness
    - Racial fairness

@@

## Metropolis-Hastings Iteration Example
To simulate redistricting, “perturbation” looks like swapping precincts along existing district borders. \ref{fig:mcmc-example} illustrates the perturbation process for an example sample in a Metropolis-Hastings chain. Perturbation works by identifying nodes that exist on congressional district borders and proposing swaps along the borders. This produces a perturbation of the previous sample which is accepted with acceptance ratio

@@

![Metropolis-Hastings iteration example (Fifield et. al.) \label{mcmc-example}](figures/mcmc_example.png)

@@

## Sources

Fifield, Benjamin, et al. "Automated Redistricting Simulation Using Markov Chain Monte Carlo." *Journal of Computational and Graphical Statistics* (2020): 1-14.

Altman, M. (1997). The computational complexity of automated redistricting: Is automation the answer. *Rutgers Computer & Tech. LJ*, 23, 81.

[`redist` Manual](https://cran.r-project.org/web/packages/redist/redist.pdf)

['sf' Manual](https://r-spatial.github.io/sf/)

[MGGG Colorado Election Shapefile](https://github.com/mggg-states/CO-shapefiles)

[Harvard Election Data Archive](https://projects.iq.harvard.edu/eda/pages/about)

https://en.wikipedia.org/wiki/Gerrymandering
